---
title: "BasicNet COVID19 Refugee Camp"
author: "Luis Chaves"
date: "18/04/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F, echo = T}
library(EpiModel)
library(parallel)
```

# Setting up empty network

## Network basic features

First, we define our population, along with their attributes.
```{r}
n = 1000 # individuals in camp

# initialize empty net
nw = network.initialize(n = n, directed = FALSE)

# set age attribute according to real age breakdown
#' Age breakdown within the camp (20,000 people)
#   50+ 4.7%      class 3
#   19-49 54.8%   class 2
#   0-18 40.5%    class 1
num.age_cl1 = round((40.5/100)*n)
num.age_cl2 = round((54.8/100)*n)
num.age_cl3 = round((4.7/100)*n)

stopifnot((num.age_cl1+num.age_cl2+num.age_cl3) == n)

nw = set.vertex.attribute(nw, "age_cl", 
                          c(rep("young", num.age_cl1),
                            rep("adults", num.age_cl2),
                            rep("old", num.age_cl3)))
```

## Network formation and dissolution parameters

Now we define the formation and dissolution terms that will be used to create out temporal network later.

In this case we'll need: 

* edges:[+1 statistic] expected # of edges in average (proxy for the mean degree)
* nodefactor("age_cl", base = 1) [+2 statistics] this will ignore the stat for class with most members (adults), this
is recommended because each tie will count twice once for each node. nodefactor() allows us to set
the expected number of edges per class/group. base = 1, sets class 1 to the base level (which I am at
this moment unsure of what it is, may be avg degree may not be)
* nodematch(age_cl, diff = FALSE):[+1 statistic] expected number of edges of individuals from the same group,
the diff flag allows you to set a single statistic that applies to everyone or you could also make
it per class.
* concurrency: [+1 statistic] number of nodes expected to have 2 or more edges at any given time

You can read more about the options that exist to customise network formation in [this paper.](http://www.jstatsoft.org/v24/i04/) as well as in the original [Epimodel methods paper](http://www.jstatsoft.org/v84/i08/).
```{r}
# formation to make network 
formation <- ~edges+nodefactor("age_cl", base = 1)+nodematch("age_cl", diff = FALSE)+concurrent

mean_degree = 2 # how many connections a person has in average
stat1 = (n/3)*mean_degree #number of expected edges per age_class in general
# these (below) need to average out to the mean degree above
mean_degree_age_cl1 = 2 # young
mean_degree_age_cl2 = 3.5 # adults
mean_degree_age_cl3 = 0.5 # old
stat2 = num.age_cl3*mean_degree_age_cl3 
stat3 = num.age_cl1*mean_degree_age_cl1
# same-race edge, let's say that only 50% of connections are between people of same age class
same_race_edge = 0.5
stat4 = stat1*same_race_edge # number of same class edges per person (see above)
# number of concurrent edges (concurrent degree), number of same connections at the same time
stat5 = n*0.5
target.stats = c(stat1,stat2, stat3, stat4, stat5)
```

Below `d.rate` is the homogeneous (average) departure rate for nodes in the network.
```{r}
coef.diss = dissolution_coefs(dissolution = ~offset(edges),
                              duration = 5, # avg duration of each node -> 100 steps
                              d.rate = 0.00525) 
coef.diss
```

Simulate network.
```{r}
# I found that the coefs are in a order that I was not expecting
est1 <- netest(nw,
               formation,
               target.stats,
               coef.diss,
               edapprox = T, # I am going to set edapprox to FALSE because the relationships
               # will last much longer than what would satisfy the conditions to use edapprox
               verbose = F)
```


```{r}
summary(est1)
```

## Network diagnostics

This step is to check everything is fine with how the newtork was set up.

```{r}
cores = detectCores()-1

t0 = Sys.time()
dx = netdx(est1,
           nsims = 5,
           nsteps = 180, # simulating 6 months
           ncores = cores)
print(paste(Sys.time()-t0, "to run the diagnostic simulations."))
dx
```

## Diagnostic plots
```{r}
plot(dx)
```

Difference between the target statistics and the estimated ones (from the simulation) may arise due to random error, in which case increasing the number of simulations will help or from systematic bias in which case the estimation approximation conditions may not hold. One reason for the estimation approximation not holding is if the dissolution coefficient is not small enough compared to the simulation length (as a rule of thumb a ratio of 1:20 between the dissolution coefficient and the simulation length should ensure a good model fit). If still a good dynamic fit is not achieved, one shoould go back to the static fit diagnostic (using \texttt{dynamic} argument in \texttt{netdx}) and if that fits then do a full STERGM estimation instead of the one using \texttt{edapprox = TRUE}.

```{r}
par(mfrow = c(1,2))
plot(dx, "duration")
plot(dx, "dissolution")
```

This results are good as the target stastitics match pretty well, could probably do more simulation for a tigther fit.

## Custom modules

We are going to make an SEIR models which is not one of the base models. Therefore we are going to need to make a new Exposed module and also change the Infection module.

### Changing infection module

Code from [this paper supplements (R replication code)](https://www.jstatsoft.org/article/view/v084i08)

The custom module arae found in another script, `NetModules.R`. 

```{r}
source("NetModules.R")
```


## Epidemic simulation

Now that we are satisfied with our diagnostics, we will model an epidemic on top of our dynamic network. First we need to set the parameters that define our pandemic. These are things such as the infection rate and infection probability upon contact of two individuals.

```{r, include = T, echo = F}
param = param.net(
                  inf.prob = 0.7,  # infection probability per transmissible act
                  # in the new infection module we have done this is actually the
                  # se. rate 
                  act.rate = 0.7, # rate of transmissible acts 
                  # per partnership per unit of time
                  ei.rate = 0.3,
                  ir.rate = 1/14, # replacces rec.rate,
                  ds.rate = 0.001,
                  de.rate = 0.005,
                  di.rate = 0.05,
                  dr.rate = 0.001,
                  a.rate = 0 # arrival rate
                  ) 

init = init.net(i.num = 50)

control = control.net(
                      # type = "SIS",
                      nsims = 2 , 
                      nsteps = 180,
                      delete.nodes = T, 
                      ncores = 1,#cores,
                      epi.by = "age_cl",
                      infection.FUN = infect,
                      progress.FUN = progress,
                      recovery.FUN = NULL,
                      skip.check = T,
                      depend = T
                      )
```


```{r,  include = F}
# so now we get a model where the total population changes as
# births and deaths are allowed
t0 = Sys.time()
sim1 = netsim(est1, param, init, control)
```


```{r}
print(Sys.time() - t0)
print("to run the network simulation.")
```

### Plot absolute numbers
```{r}
plot(sim1, y = c("num", "r.num", "e.num", "i.num", "s.num"), legend = T) # type epi is the default
```

### Plot flows
```{r}
plot(sim1, y  = c("se.flow","ei.flow", "ir.flow"), legend = T)
```

### Plots by categories
```{r}
plot(sim1, y = c("i.num.age_clyoung", "i.num.age_cladults", "i.num.age_clold"),
     legend = T)
```

### Plot number by category
```{r}
plot(sim1, y = c("num.age_clyoung", "num.age_clold", "num.age_cladults", "num"),
     legend = T, grid = T)

```

### Plot deaths

Flows are not super useful
```{r}
plot(sim1, y = c("de.flow", "dr.flow", "ds.flow", "di.flow"), legend = T)
```


```{r}
# plotting options
#?plot.netsim
#plot(sim1, "network", at = 1, sims = "mean", col.status = T, legend = T)
# plot(sim1, "network", at = 500, sims = "mean", col.status = T, legend = T)
```

Retrieve the whole dataframe for further analysis.
```{r}
head(as.data.frame(sim1))
```




