---
title: "SEIQRHF Network Model"
author: "Luis Chaves"
output:
  html_document:
    df_print: paged
always_allow_html: yes
---

```{r doc_options, echo = F, message = F}
knitr::opts_chunk$set(echo = T, warning = F, message  = F)
```


```{r packages}

library(tidyverse) # useful for general R
library(EpiModel) # EpiModel package
library(parallel) # for parallel processing where applicable
library(plotly) # for interactive plots
library(foreach)
theme_set(theme_bw())

if("igraph" %in% (.packages())){ # igraph sometimes clashes with the statnet and the network package
  detach("package:igraph", unload=TRUE) 
}

calculate.max.edges = function(n){
  return (n*(n-1)/2)
}
```

# Modelling COVID-19 spread in the Moria refugee camp

This model builds on previous knowledge acquired through weeks of work and therefore does not aim to be an instructive piece for beginers to EpiModel. For the old notebooks that were used to build knowledge please see Archive/DevNotebooks00.

## Population structure & attributes

### Read in data

At this time we have individual information on age, gender and aggregate data on ethnicity.
```{r}
# age and gender data
ageData = read.csv("../Data/age_and_sex.csv") %>% rename(Age = V1, Gender = V2, ID = X)

# camp parameters: hospitalisation rate and fatality rate by Age group, from Nick in AI4GOOD team
campParams = read.csv("../Data/camp_params.csv") %>% rename(Pop_Proportion = Value) %>%
  filter(Camp  == "Camp_1", Variable == "Population_structure")
campParams$Age = gdata::drop.levels(campParams$Age)

# processing of the age data to be put into 10 year age bins
ageGroups = campParams %>%
  select(Age) %>% as.matrix()

# left inclusive and right exclusive cuts
ageData$ageGroup = cut(ageData$Age, breaks = c(0,10,20,30,40,50,60,70,80, Inf))
```

### Visualise age distributions and risk by age group

```{r}
plot(as.factor(campParams$Age), campParams$Hosp_given_symptomatic,
     main = "Hospitalisation rate by age group",
     ylab = 'Hospitalisation rate', xlab = 'Age group')

plot(as.factor(campParams$Age), campParams$Critical_given_hospitalised,
     main = "Fatality rate by age group",
     ylab = 'Fatality rate', xlab = 'Age group')

plot(density(ageData$Age), main = "Distribution of age in Moria camp",
     xlab = 'Age', ylab = 'Density')
```

### Put parameters into data frame
```{r}
paramsFromData = list()
paramsFromData$age.dist = ageData$ageGroup
# the two below are in same order as age groups
paramsFromData$rates.byAge = data.frame(AgeGroup = levels(ageData$ageGroup),
                                        hosp.rate = campParams$Hosp_given_symptomatic,
                                        fat.rate = campParams$Critical_given_hospitalised)
```

## Network structure based on housing allocation

Individuals in Moria live in either tents or isoboxes, with a capacity of 4 and 10 respectively. In this model we attribute each person to a housing unit which is represented by a unique name (e.g. tent1, tent2, isobox4...). From the Tucker model, we know that 8100 indviduals occupy isoboxes and 10600 are in tents (total of 18700). We can also use this numbers to make proportions by housing unit and build smaller networks (less individuals).

_Note:_ For convenience, we will define here `n`, the number of individuals in the simulated network. In experiments, this number equals to number of individuals in Moria.

```{r}
n = 1000
```


```{r}
prop.isobox = round(8100/(8100+10600),2) # prop. of people in isobox
prop.tent = 1 - prop.isobox # prop. of people in tent

cat(paste0("Proportion of people in isoboxes: ", prop.isobox, "\n",
           "Proportion of people in tents:", prop.tent, "\n"))
```

### Making the housing attribute vectors

The number of tents and the number of isoboxes are calculated as:
$$N_{tents} = \frac{\text{Number of people in tents}}{\text{tent capacity}} = \frac{n\cdot Prop_{tents}}{tent_{capacity}}\\  N_{isobox} = \frac{\text{Number of people in isoboxes}}{\text{isobox capacity}}= \frac{n\cdot Prop_{isobox}}{isobox_{capacity}}$$

_Note_: this are rounded to get integer values.

```{r}
tent.capacity = 4
iso.capacity = 10

num_of_iso = round(n*prop.isobox/iso.capacity) # number of isoboxes
num_of_tents = round(n*prop.tent/tent.capacity) # number of tents

if (n > num_of_iso*iso.capacity+num_of_tents*tent.capacity){
  num_of_tents = num_of_tents+1
}

num_in_iso = num_of_iso*iso.capacity
num_in_tents = n - num_in_iso # num_of_tents*tent.capacity 

cat(paste0(
  "Number of tents: ", num_of_tents, "\n",
  "Number of isoboxes: ", num_of_iso, "\n",
  "Number of people in tents: ", num_in_tents, "\n",
  "Number of people in isoboxes: ", num_in_iso, "\n\n",
  "Number of housing units in total: ", num_of_iso+num_of_tents))
```

## Allocating people to rooms
```{r}
iso_ids = 1:num_of_iso
tent_ids = 1:num_of_tents 


housing_iso = apportion_lr(
  vector.length = num_in_iso,
  values = iso_ids,
  proportions = rep(1/(num_of_iso), num_of_iso)
)
housing_iso = paste0("iso", housing_iso)


housing_tents = apportion_lr(
  vector.length = num_in_tents,
  values = tent_ids,
  proportions = rep(1/(num_of_tents), num_of_tents)
)
housing_tents = paste0("tent", housing_tents)


housing = c(housing_iso,housing_tents)

# residence vector to keep track of those in tents and isos
residence = c(rep("iso", num_in_iso), rep("tent", num_in_tents))

plot(table(housing),
     main = "Number of people per housing unit",
     ylab = "Number of people",
     xlab = "Housing ID",
)
grid()

```

## Creating network and setting node attributes

```{r}
nw = network.initialize(n = n, directed = FALSE)

# housing
nw = set.vertex.attribute(nw, "housing", housing)
# age
nw = set.vertex.attribute(nw, "age", sample(as.vector(paramsFromData$age.dist),n))
```

## Setting edge formation dynamics
```{r}
formation = ~edges + 
  offset(nodematch("housing", diff = F))

## max in-house ties
max.inhouse.edges = 0
for (num_in_house in table(housing)){
  max.inhouse.edges = max.inhouse.edges + calculate.max.edges(num_in_house)
}

# default degrees in housing units as per current occupancy
## The degree of a node in a network is the number of connections it has to other nodes
# the -1 is tou account fr the lack of connection to one-self
iso.default.degree = mean(table(housing)[iso_ids]-1)
tent.default.degree = mean(table(housing)[tent_ids]-1)

# number of external contacts per person in average
external.contacts = 4

mean_degree.iso =  iso.default.degree + external.contacts
mean_degree.tent =  tent.default.degree + external.contacts

# calculate mean degree
mean_degree = (num_in_iso*mean_degree.iso+
                 num_in_tents*mean_degree.tent)/n

expected.edges = n*mean_degree/2

target.stats = c(expected.edges)

d.rate = 0
coef.diss = dissolution_coefs(dissolution = ~offset(edges)+
                                offset(nodematch("housing", diff = F)),
                              duration = c(2, 1e9),
                              d.rate = d.rate)
```

# Network estimation step
```{r}
est <- netest(nw,
              formation,
              target.stats,
              coef.diss,
              coef.form = Inf,
              set.control.ergm = control.ergm(MCMLE.maxit = 500)
)
summary(est)
# mcmc.diagnostics(est1$fit)
```

### Network diagnostics

Do our network simulation match our target statistics?
```{r}
cores = parallel::detectCores()-1

dx <- netdx(est,
            nsims = 1e3,
            nsteps = 90,
            ncores = cores,
            dynamic = FALSE,
            nwstats.formula = ~edges + nodematch("housing", diff = FALSE),
            set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e6),
            keep.tnetwork = T)

print(dx) # diagnostics table

plot(dx) # diagnostics plot

```

## Simulating epidemic in our dynamic network

First we import our custom modules
```{r}
source("../Scripts/SEIQHRFNetModules.R")
```

Next we set our epidemic parameters
```{r}
param = param.net(act.rate.se = 10,
                  inf.prob.se = 0.02,
                  act.rate.si = 10,
                  inf.prob.si = 0.05,
                  act.rate.sq = 2.5,
                  inf.prob.sq = 0.02,
                  ei.rate = 1/10,
                  iq.rate = 1/30, #c(rep(1/30, 60), rep(15/30, 120)), # time varying works
                  ih.rate = 1/100,
                  qh.rate = 1/100,
                  hr.rate = 1/15,
                  qr.rate = 1/20,
                  hf.rate = 1/50,
                  hf.rate.overcap = 1/25,
                  hosp.cap = 5,
                  hosp.tcoeff = 0.5,
                  a.rate = 0,
                  di.rate = d.rate,
                  ds.rate = d.rate,
                  dr.rate = d.rate,
                  ratesbyAge = paramsFromData$rates.byAge
) 

init = init.net(i.num = 1,
                r.num = 0,
                e.num = 0,
                s.num = n - 1,
                f.num = 0,
                h.num = 0,
                q.num = 0
)
```


```{r echo = F, message = F}
control = control.net(
  nsims = 3, 
  nsteps = 90,
  # delete.nodes = T,  this does not work for now
  ncores = cores,
  initialize.FUN = custom.initialize.net, # this bit is just so that I can extract time
  exposure.FUN = exposure,
  infect.FUN = infect,
  epi.by = "age",
  quarantine.FUN = quarantining,
  hospitalize.FUN = RequireHospitalization,
  recover.FUN = recover,
  fatality.FUN = fatality,
  recovery.FUN = NULL,
  infection.FUN = NULL,
  departures.FUN = departures.net,
  get_prev.FUN = custom.get_prev.net,
  skip_check = FALSE,
  depend = T
)

t0 = Sys.time()
sim = custom.netsim(est, param, init, control)
```


```{r}
print(Sys.time()-t0) # roughly 30 minutes for n = 1000 and nsteps = 60
res = as.data.frame(sim)
```

## Some more network diagnostic

```{r}
nw_sim <- get_network(sim)

housing <- nw_sim %v% "housing"
summary(res$duration[which(housing[res$head] == housing[res$tail])])
summary(res$duration[which(housing[res$head] != housing[res$tail])])
```



## Results visualisation

### Overall numbers for each tracked state
```{r}
ggplotly(
  res %>% select(s.num, e.num, i.num, q.num, h.num, r.num, f.num, num, time) %>%
    group_by(time) %>% summarise_all(~mean(.)) %>% 
    pivot_longer(-time) %>% ggplot(aes(x = time, y = value, color = name))+
    geom_line(size = 1)+scale_color_brewer(palette = "Set1")
)
```

### Characteristic curves
```{r}
get_times <- function(simulation.object) {
  
  sim <- simulation.object
  
  for (s in 1:sim$control$nsims) {
    if (s == 1) {
      times <- sim$times[[paste0("sim", s)]]
      times <- times %>% mutate(s = s)
    } else {
      times <- times %>% bind_rows(sim$times[[paste("sim", 
                                                    s, sep = "")]] %>% mutate(s = s))
    }
  }
  
  times <- times %>%
    mutate(infTime = ifelse(infTime < 0, -5, infTime),
           expTime = ifelse(expTime < 0, -5, expTime)) %>% 
    mutate(incubation_period = infTime - expTime,
           illness_duration = recTime - expTime,
           illness_duration_hosp = dischTime - expTime, 
           hosp_los = dischTime - hospTime,
           quarantine_delay = quarTime - infTime,
           survival_time = fatTime - infTime) %>% 
    select(s,
           incubation_period,
           quarantine_delay,
           illness_duration, 
           illness_duration_hosp,
           hosp_los,
           survival_time) %>% 
    pivot_longer(-s, names_to = "period_type", values_to = "duration") %>% 
    mutate(period_type = factor(period_type,
                                levels = c("incubation_period", 
                                           "quarantine_delay",
                                           "illness_duration",
                                           "illness_duration_hosp", 
                                           "hosp_los",
                                           "survival_time"),
                                labels = c("Incubation period", 
                                           "Delay entering isolation",
                                           "Illness duration",
                                           "Illness duration (hosp)", 
                                           "Hospital care required duration",
                                           "Survival time of case fatalities"), 
                                ordered = TRUE))
  return(times)
}
```

```{r}
times = get_times(sim)

times %>% filter(duration <= 30) %>% ggplot(aes(x = duration)) + 
  geom_density() + facet_wrap(period_type ~ ., scales = "free_y") + 
  labs(title = "Duration frequency distributions", subtitle = "Baseline simulation")

```

## Network visualisation

As a heatmap
```{r}
nw_object = get_network(sim)
net_at_1 = network.collapse(nw_object, at = 2)

graph = intergraph::asIgraph(net_at_1)

adj = igraph::as_adjacency_matrix(graph, sparse = F)
colnames(adj) = igraph::V(graph)$housing
rownames(adj) = igraph::V(graph)$housing

adj = adj[order(rownames(adj)), order(colnames(adj))]

pheatmap::pheatmap(adj,
                   color = c("grey50","black"),
                   border_color = "white",
                   angle_col = 45,
                   angle_row = 45,
                   fontsize = 6,
                   legend_breaks = c(0,1),
                   legend = F,
                   cluster_rows = F,
                   cluster_cols = F,
                   show_rownames = ifelse(n<100, T, F),
                   show_colnames = ifelse(n<100, T, F))

```

## Network heatmap GIF

```{r, animation.hook="gifski"}
get_animation = T
if (get_animation) {
  library(animation)
  
  nw_object = get_network(sim)
  ani.record(reset = TRUE)  # clear history before recording
  for (at in c(1:30)){
    
    net_at = network.collapse(nw_object, at = at)
    graph = intergraph::asIgraph(net_at)
    adj = igraph::as_adjacency_matrix(graph, sparse = F)
    
    colnames(adj) = igraph::V(graph)$housing
    rownames(adj) = igraph::V(graph)$housing
    
    adj = adj[order(rownames(adj)), order(colnames(adj))]
    
    pheatmap::pheatmap(adj,
                       color = c("grey50","black"),
                       border_color = "white",
                       angle_col = 45,
                       angle_row = 45,
                       fontsize = 6,
                       legend_breaks = c(0,1),
                       legend = F,
                       cluster_rows = F,
                       cluster_cols = F,
                       show_rownames = ifelse(n<100, T, F),
                       show_colnames = ifelse(n<100, T, F))
    ani.record()
  }
  
  oopts = ani.options(interval = 0.5)
}

```


```{r }
# ani.replay()
# saveHTML(ani.replay(), img.name = "record_plot")

```

